---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
# Install missing packages
list.of.packages <- c("ggplot2", "data.table", "xtable")
new.packages <- list.of.packages[!(list.of.packages %in%     
  installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

unzip(zipfile = 'activity.zip')
activity_df <- read.csv(file = 'activity.csv')
```


## What is mean total number of steps taken per day?
```{r}
library(data.table)
steps_by_day <- setDT(activity_df[!is.na(activity_df$steps), ])[, list(steps_daily=sum(steps)), by = 'date']
```
```{r, results='asis'}
library(xtable)
xt <- xtable(head(steps_by_day))
print(xt, type='html')
```

Histogram of the total number of steps taken each day:
```{r}
library(ggplot2)
qplot(steps_daily, data=steps_by_day, bins=20, fill = I('darkgreen'))
```

Mean of the total number of steps taken per day:
```{r}
mean(steps_by_day$steps_daily, na.rm = TRUE)
```
Median of the total number of steps taken per day:
```{r}
median(steps_by_day$steps_daily, na.rm = TRUE)
```




## What is the average daily activity pattern?
```{r}
activity_pattern <- setDT(activity_df[!is.na(activity_df$steps), ])[, list(mean_steps_in_interval=mean(steps)), by = 'interval']
max_mean_steps <- activity_pattern$interval[which.max(activity_pattern$mean_steps_in_interval)]
ggplot(activity_pattern, aes(interval, mean_steps_in_interval)) + geom_line() + geom_vline(xintercept = max_mean_steps, col='red', lwd=1)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
activity_pattern$interval[which.max(activity_pattern$mean_steps_in_interval)]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity_df))
```
Imputing strategy: use the mean for a given interval across all days.
```{r}
activity_df_imputed <- activity_df
missing_idx <- which(is.na(activity_df$steps))
activity_df_imputed$steps[missing_idx] <- sapply(missing_idx, function(idx) activity_pattern$mean_steps_in_interval[which(activity_pattern$interval==activity_df$interval[idx])] )
length(which(is.na(activity_df_imputed$steps))) # Verify that no missing values remained
```
Histogram of the total number of steps taken each day
```{r}
steps_by_day_imputed <- setDT(activity_df_imputed)[, list(steps_daily=sum(steps)), by = 'date']
qplot(steps_daily, data=steps_by_day_imputed, bins=20, fill = I('darkgreen'))
```

Mean of the total number of steps taken per day:
```{r}
mean(steps_by_day_imputed$steps_daily, na.rm = TRUE)
```
Median of the total number of steps taken per day:
```{r}
median(steps_by_day_imputed$steps_daily, na.rm = TRUE)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale("LC_TIME","C")
activity_df_imputed$is_weekend <- factor(x=ifelse(weekdays(as.Date(activity_df_imputed$date)) %in% c('Saturday', 'Sunday'),'weekend', 'weekday'))
# g <- ggplot(activity_df_imputed, aes(interval, steps)) + facet_grid()
activity_pattern_imputed <- setDT(activity_df_imputed)[, list(mean_steps_in_interval=mean(steps)), by = c('interval', 'is_weekend') ]
qplot(interval, mean_steps_in_interval, data=activity_pattern_imputed, facets = is_weekend~., geom='line', ylab='Number of steps', xlab='Interval', color=I('blue'))
```
